### 2014-04-07 doc/thesis/introduction.tex
Minor text adaptations to the introduction

Dynamically typed languages such as Lua avoid static types in favor of
simplicity and flexibility, because the absence of static types means
that programmers do not need to bother about abstracting types that
should be validated by a type checker.
Instead, dynamically typed languages use run-time \textit{type tags}
to classify the values they compute, so their implementation can use
these tags to perform run-time (or dynamic) type checking
\citep{pierce2002tpl}.

@@

Ideally, programming languages should offer programmers the
option to choose between static and dynamic typing:
\textit{optional type systems} \citep{bracha2004pluggable} and
\textit{gradual typing} \citep{siek2006gradual} are two similar
approaches for blending static and dynamic typing in the same
language.
The aim of both approaches is to offer programmers the option
to use type annotations where static typing is needed and avoid them
where dynamic typing is sufficient.
The difference between these two approaches is the way they treat
run-time semantics.
While optional type systems do not affect the run-time semantics,
gradual typing uses run-time checks to ensure that dynamically typed
code does not violate the invariants of statically typed code.

@@

In this work we propose the design and implementation of Typed Lua:
an optional type system for Lua that should be complex enough to
preserve some of the Lua idioms that programmers are already used to,
but that should also include new constructs that help programmers
structure Lua programs.

@@

So far, Typed Lua should be a strict superset of Lua that
provides optional type annotations, compile-time type checking, and
class-based object-oriented programming through the definition of
classes, interfaces, and modules.
In practice, we should implement Typed Lua as a programming
language that extends Lua syntax to add optional type annotations
and standard constructions for structuring Lua code.
The compiler should use static types to perform compile-time type
checking, but should also allow Lua code to coexist with Typed Lua
code, and generate Lua code that runs in unmodified Lua
implementations.
We have an ongoing implementation of the Typed Lua compiler that is
available online\footnote{https://github.com/andremm/typedlua}.

@@

We also believe that Typed Lua will help programmers give a more
formal documentation to already existing Lua code, as static types
are also a useful source of documentation in languages that provide
type annotations, because type annotations are always validated by
the type checker and therefore never get outdated.
Thus, programmers can use Typed Lua to define axioms about the
interfaces and types of dynamically typed modules.
We will enforce this point by using Typed Lua to statically type
the interface of the Lua standard libraries and other commonly used
Lua libraries, so our compiler can check their usage by Typed Lua
code.

Typed Lua will perform a very limited form of local type inference
\citep{pierce2000lti}, as static typing does not necessarily mean
that programmers need to insert type annotations in the code.
Several statically typed languages such as Haskell provide some
amount of type inference that automatically deduces the types of
expressions.
Still, Typed Lua will need a small amount of type annotations due
to the nature of its optional type system.

Typed Lua will not deal with code optimization, although another
important advantage of static types is that they help the compiler
perform optimizations and generate more efficient code.
However, we believe that the formalization of our optional type
system will be precise enough to aid optimization in some Lua
implementations.

We will use some of the ideas of gradual typing to formalize Typed Lua.
Even though Typed Lua is an optional type system and thus will not
include run-time checks between dynamic and static regions of the
code, we believe that using the foundations of gradual typing to
formalize our optional type system will allow us to include run-time
checks in the future.

%In Section \ref{sec:review} we review the literature about static and
%dynamic typing in the same language, which includes the discussion
%about the particularities between optional type systems and gradual
%typing.
%In Section \ref{sec:proposal} we describe the reasons why Typed Lua
%includes features of object-oriented programming and the status of
%its design and implementation.
%In Section \ref{sec:con} we outline our expected contributions.
%In Section \ref{sec:plan} we present the work plan for thesis completion.

>>>>

Dynamically typed languages such as Lua avoid static types in favor of
simplicity and flexibility, because the absence of static types means
that programmers do not need to bother about abstracting types that
should be validated by a type checker.
Instead, dynamically typed languages use run-time \emph{type tags}
to classify the values they compute, so their implementation can use
these tags to perform run-time (or dynamic) type checking
\citep{pierce2002tpl}.

@@

Ideally, programming languages should offer programmers the
option to choose between static and dynamic typing:
\emph{optional type systems} \citep{bracha2004pluggable} and
\emph{gradual typing} \citep{siek2006gradual} are two similar
approaches for blending static and dynamic typing in the same
language.
The aim of both approaches is to offer programmers the option
to use type annotations where static typing is needed and to not use
them where dynamic typing is sufficient.
The difference between these two approaches is the way they treat
run-time semantics.
While optional type systems do not affect the run-time semantics,
gradual typing uses run-time checks to ensure that dynamically typed
code does not violate the invariants of statically typed code.

@@

In this work we present the design and implementation of Typed Lua:
an optional type system for Lua that is complex enough to
preserve some of the Lua idioms that programmers are already used to,
but that also includes new constructs that help programmers
structure Lua programs.

@@

So far, Typed Lua is a strict superset of Lua that
provides optional type annotations, compile-time type checking, and
class-based object-oriented programming through the definition of
classes, interfaces, and modules.
In practice, we implement Typed Lua as a programming language that
extends Lua syntax to add optional type annotations and standard
constructions for structuring Lua code.
The compiler uses static types to perform compile-time type
checking, but also allows Lua code to coexist with Typed Lua code,
and generates Lua code that runs in unmodified Lua implementations.
We have an ongoing implementation of the Typed Lua compiler that is
available online\footnote{https://github.com/andremm/typedlua}.

@@

We also believe that Typed Lua will help programmers give a more
formal documentation to already existing Lua code, as static types
are also a useful source of documentation in languages that provide
type annotations, because type annotations are always validated by
the type checker and therefore never get outdated.
Thus, programmers can use Typed Lua to define axioms about the
interfaces and types of dynamically typed modules.
We enforce this point by using Typed Lua to statically type
the interface of the Lua standard libraries and other commonly used
Lua libraries, so our compiler can check their usage by Typed Lua
code.

Typed Lua performs a very limited form of local type inference
\citep{pierce2000lti}, as static typing does not necessarily mean
that programmers need to insert type annotations in the code.
Several statically typed languages such as Haskell provide some
amount of type inference that automatically deduces the types of
expressions.
Still, Typed Lua needs a small amount of type annotations due
to the nature of its optional type system.

Typed Lua does not deal with code optimization, although another
important advantage of static types is that they help the compiler
perform optimizations and generate more efficient code.
However, we believe that the formalization of our optional type
system is precise enough to aid optimization in some Lua implementations.

We use some of the ideas of gradual typing to formalize Typed Lua.
Even though Typed Lua is an optional type system and thus does not
include run-time checks between dynamic and static regions of the
code, we believe that using the foundations of gradual typing to
formalize our optional type system will allow us to include run-time
checks in the future.

In Chapter \ref{chap:review} we review the literature about static and
dynamic typing in the same language, which includes the discussion
about the particularities between optional type systems and gradual
typing.
In Chapter \ref{chap:typedlua} we present the design and implementation
of Typed Lua along with the reasons why it includes features of
object-oriented programming.
In Chapter \ref{chap:cases} we present some case studies.
In Chapter \ref{chap:conc} we outline our contributions.
